The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-13 paper.tex
More or less finished Chaha, added stringency info and Bangla analysis to mechanisms section

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[table]{xcolor}    %for shading*
\usepackage{fullpage}
\usepackage{natbib}
\usepackage{mathpazo}
\usepackage{tipa}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{linguex}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{slashbox}	
\usepackage{pifont}    %for pointing hand
\usepackage{arydshln}    %for dashed lines
\usepackage{rotating}    %for angled text
\usepackage{wasysym}  % for frowny face

@@

This faithfulness-based approach raises the question of how the perceptual distance between two sonority contours /C\textsubscript{1}C\textsubscript{2}/ and [C\textsubscript{1}VC\textsubscript{2}] should be computed in terms of the sonority of C\textsubscript{1}, C\textsubscript{2} and V.  
Fleischhacker's analysis rested on empirical determinations of sonority contour faithfulness, and did not attempt to determine such a relation.

\citet{steriade.2006} proposed that input and output sonority contours should match in terms of whether they are rising or falling, and to what degree, but did not suggest a concrete mathematical relation.  \citet{flemming.2008} formalises Steriade's approach with the metric {\sc Sonority Rise}, the ratio of the gradients of the two contours.

@@

In this paper, I suggest an alternative metric, {\sc Sonority Angle}, namely the magnitude of the angle made by the vectors C\textsubscript{1}-C\textsubscript{2} and C\textsubscript{1}-V, and explore the ramifications of this choice.

{\sc Sonority Angle} makes the same broad predictions as {\sc Sonority Rise} -- that clusters of rising sonority, having a relatively small angle between the underlying sonority contour /C\textsubscript{1}-C\textsubscript{2}/ and the overt sonority contour [C\textsubscript{1}-V], are perceptually more similar to their epenthetic output, and therefore more likely to undergo epenthesis, than clusters of falling sonority. Crucially, however, the exact hierarchy of susceptibility of individual clusters to epenthesis is predicted to be different.

@@

Syllable Contact forms the basis for the main existing analysis of Chaha epenthesis by \citet{rose.2000}.  I show that the faithfulness-based analysis, powered by the metric of {\sc Sonority Angle}, is more economical.  In the case of Irish, Syllable Contact makes incorrect predictions regarding the data.

@@

The second generalisation is that the more sonorous the C\textsubscript{1}, the more likely
epenthesis is to occur. This time, fix C\textsubscript{2} and lower the sonority of C\textsubscript{1}.

\ex. \begin{tikzpicture} [shorten >=1pt,scale=0.33]
                      \draw [<-] (-3,6.5) -- (-3, 0.5) ; % axis
                      \node at (-3, 7.0) {Sonority}; % axis label
					  % left diagram
                      \draw (5,1) -- (0,3) -- (5,6) ; % rising sonority
                      \node at (1.8, 3.2) {$\theta$}; % theta label for rising sonority
                      \node[left] at  (0,3) {C$_1$}; 
                      \node[right] at (5,1) {C$_2$};
                      \node[right] at (5,6) {V};
					  % right diagram
                      \draw (20,1) -- (15,1) -- (20,6) ; % falling sonority
                      \node at (17, 1.9) {$\theta$}; % theta label for falling sonority
                      \node[left] at (15,1) {C$_1$}; 
                      \node[right] at (20,1) {C$_2$};
                      \node[right] at (20,6) {V};
\end{tikzpicture}

While the difference is less clear visually, the second angle is larger. The first term in the formula confirms the relation between the sonority of C\textsubscript{1} in terms of its closeness to V, and {\sc Sonority Angle} as a whole.

@@

These faithfulness constraints have a universal ranking, with the least stringent the highest-ranked.

   \ex. ... \newline
            $\gg$ \textsc{Ident(SonAngle)}$<$1.5 \newline
            $\gg$ \textsc{Ident(SonAngle)}$<$1.0 \newline
            $\gg$ \textsc{Ident(SonAngle)}$<$0.5 \newline
            $\gg$ ...

The resulting hierarchy of clusters, ranked according to their resistance to epenthesis as defined
by their {\sc Sonority Angle}, is as follows.

@@

Notice that out of the falling sonority clusters, those that decrease in sonority by a single step
-- namely RN, NF and FT -- have smaller {\sc Sonority Angles} than the ones that have a greater fall
in sonority. Furthermore, out of these three clusters, theone with the most sonorous C\textsubscript{1}, RN, has the smallest {\sc Sonority Angle}. We thus predict that out of the falling sonority clusters,
RN is the most likely to be broken up by epenthesis. The case study on Chaha will demonstrate that this
is the case.

@@

@@

I will contrast {\sc Sonority Angle} with an alternative metric of sonority contour faithfulness proposed by \citet{flemming.2008}, {\sc Sonority Rise}, which takes the ratio of the underlying sonority contour /C\textsubscript{1}C\textsubscript{2}/ and the surface contour [C\textsubscript{1}V].

@@

{\sc Sonority Rise} also gives rise to a hierarchy of susceptbility of clusters.

@@

Though in many ways similar to the {\sc Sonority Angle} hierarchy, it makes several crucially different predictions. For example, it does not share the prediction of {\sc Sonority Angle} that RN is the most likely of the falling sonority clusters to epenthesise -- rather, if RN undergoes epenthesis then we expect FT and NF to do the same. Furthermore, it predicts that if NT and RT fail to undergo epenthesis due to the high faithfulness cost of interrupting these clusters, then RF should also fail to undergo epenthesis. Both of these predictions are contrary to the evidence of Irish and Chaha.

@@

\ex. Universal ranking cannot derive [\textipa{j@ms1to}]

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-mst-o}/} & {\sc Ident(SonAngle)}$<$2.1 & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a.  &\frownie & \textipa{j@ms1to} & 2.11 $\ast$! & \cellcolor{lightgray}2.11 $\ast$ & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@m1sto} & 2.03 & \cellcolor{lightgray}2.03 $\ast$ & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

@@

\ex. TODO tableaux

@@

\ex. TODO tableaux

@@

\subsection{Comparison with alternative approaches}

>>>>

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[table]{xcolor}    %for shading*
\usepackage{fullpage}
\usepackage{natbib}
\usepackage{mathpazo}
\usepackage{tipa}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{linguex}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage{slashbox}	
\usepackage{pifont}    %for pointing hand
\usepackage{arydshln}  %for dashed lines
\usepackage{rotating}  %for angled text
\usepackage{wasysym}   %for frowny face

@@

This faithfulness-based approach raises the question of how the perceptual distance between two sonority contours /C\textsubscript{1}C\textsubscript{2}/ and [C\textsubscript{1}VC\textsubscript{2}] should be computed in terms of the sonority of C\textsubscript{1}, C\textsubscript{2} and V.

Fleischhacker's analysis rested on empirical determinations of sonority contour faithfulness, and did not attempt to determine such a relation. \citet{steriade.2006} proposed that input and output sonority contours should match in terms of whether they are rising or falling, and to what degree, but did not suggest a concrete mathematical relation.  \citet{flemming.2008} formalises Steriade's approach with the metric {\sc Sonority Rise}, the ratio of the gradients of the two contours.

@@

In this paper, I suggest that any such metric should bear two properties. The first is that suggested by previous analyses, that the more positive the gradient of the sonority contour from C1 to C2, the smaller the faithfulness cost and hence the more likely epenthesis is to occur.
The second is that for a given sonority distance between C1 and C2, the more sonorous C1 and C2 are, the more likely epenthesis is to occur.

To make things concrete, I suggest a metric that possesses these two properties, which I term {\sc Sonority Angle}. This is the magnitude of the angle made by the vectors C\textsubscript{1}-C\textsubscript{2} and C\textsubscript{1}-V. I then explore the ramifications of this choice.

{\sc Sonority Angle} makes the same broad predictions as {\sc Sonority Rise} --- that clusters of rising sonority, having a relatively small angle between the underlying sonority contour /C\textsubscript{1}-C\textsubscript{2}/ and the overt sonority contour [C\textsubscript{1}-V], are perceptually more similar to their epenthetic output, and therefore more likely to undergo epenthesis, than clusters of falling sonority. Crucially, however, the exact hierarchy of susceptibility of individual clusters to epenthesis is predicted to be different.

@@

Syllable Contact forms the basis for the main existing analysis of Chaha epenthesis by \citet{rose.2000}.  I show that the faithfulness-based analysis, powered by the metric of {\sc Sonority Angle}, is slightly more economical in addressing the data. Furthermore, in the case of Irish, Syllable Contact makes incorrect predictions regarding the data.

@@

The second generalisation is that the more sonorous the C\textsubscript{1}, the more likely
epenthesis is to occur. Let us fix the sonority distance between the two consonants of the cluster at -2, but vary their distance from V: %TODO : check if it should be -2 or +2

\ex. \begin{tikzpicture} [shorten >=1pt,scale=0.33]
                      \draw [<-] (-3,6.5) -- (-3, 0.5) ; % axis
                      \node at (-3, 7.0) {Sonority}; % axis label
					  % left diagram
                      \draw (5,1) -- (0,3) -- (5,6) ; % 
                      \node at (1.8, 3.2) {$\theta$}; % 
                      \node[left] at  (0,3) {C$_1$}; 
                      \node[right] at (5,1) {C$_2$};
                      \node[right] at (5,6) {V};
					  % right diagram
                      \draw (20,3) -- (15,5) -- (20,6) ; % rising sonority
                      \node at (18, 4.8) {$\theta$}; % theta label for falling sonority
                      \node[left] at (15,5) {C$_1$}; 
                      \node[right] at (20,3) {C$_2$};
                      \node[right] at (20,6) {V};
\end{tikzpicture}

When C1 (and, by extension, C2) is of higher sonority, as on the right, the V-C1-C2 angle is clearlysmaller. The first term in the formula confirms the relation between the sonority of C\textsubscript{1} in terms of its closeness to V, and {\sc Sonority Angle} as a whole.

@@

This family of faithfulness constraints forms a stringency hierarchy. Any cluster that would, upon epenthesis, violate \textsc{Ident(SonAngle)}$<$n would also violate {\sc Ident(SonAngle)}$<$m where m$>$n.
The resulting hierarchy of clusters, ranked according to their resistance to epenthesis as defined
by their {\sc Sonority Angle}, is as follows.

@@

The prediction is that when a markedness constraint targets a set of clusters for epenthesis, if a certain cluster undergoes epenthesis, then all clusters to the left of it on the hierarchy --- those with a smaller {\sc Sonority Angle} --- should also undergo epenthesis.

Notice that out of the falling sonority clusters, those that decrease in sonority by a single step
-- namely RN, NF and FT -- have smaller {\sc Sonority Angles} than the ones that have a greater fall
in sonority. Furthermore, out of these three clusters, the one with the most sonorous C\textsubscript{1}, RN, has the smallest {\sc Sonority Angle}. We thus predict that out of the falling sonority clusters, RN is the most likely to be broken up by epenthesis. The case study on Chaha will demonstrate that this is the case.

@@

\subsection{Stringency and faithfulness conflation}

There are two possible ways to rank the family of {\sc Ident(SonAngle)}$<$n constraints.
One is to fix them in a universal ranking \citep{prince.smolensky.1993}, with least stringent the highest-ranked:

\ex. Universal ranking: \\
     {\sc Ident(SonAngle)}$<$2.0 $\gg$ {\sc Ident(SonAngle)}$<$1.5 $\gg$ ...

It is also possible to allow them to be freely ranked \citep[and others]{de.lacy.2004}, so that a more stringent constraint may outrank a less stringent one:

\ex. Possible free ranking:
     {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Ident(SonAngle)}$<$2.0 $\gg$ ...

Depending on what constraints intervene, this may result in conflation of certain areas of the hierarchy, so that any distinctions between {\sc Sonority Angles} between 1.5 and 2.0 may be conflated.

In this paper, for ease of analysis, I will assume that all {\sc Ident(SonAngle)}$<$n constraints are ranked low by default, but that individual {\sc Ident(SonAngle)}$<$n constraints can be promoted to dominate a markedness constraint $\mathcal{M}$ against clusters or another faithfulness constraint $\mathcal{F}$, blocking epenthesis for all clusters with a {\sc Sonority Angle} greater than $n$, with the result that either the cluster remains intact, or a competing solution to epenthesis emerges.

An example of {\sc Ident(SonAngle)}$<$n blocking epenthesis is given in the sample tableaux below.

\ex. Ident(SonAngle)$<$1.5 $\gg$ {\sc *ComplexCoda} $\gg$ {\sc Dep}
     \a. Allows epenthesis in rising and level clusters such as stop-fricative TF:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline a. & & TF & & $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Prevents epenthesis in falling clusters such as FT:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc *ComplexCoda} & {\sc Dep} \\[0.5ex]
\hline \hline c. & \ding{43} & FT & &\cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} \\
\hline d. & & F\textipa{@}T & 2.11 $\ast$! &\cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

We will see more examples of this in the analysis of Irish epenthesis as well as Chaha coda cluster epenthesis.

I will illustrate the use of {\sc Ident(SonAngle)}$<$n to block epenthesis in favour of other cluster resolution strategies with an example from the loanword phonology of the Dhaka dialect of Bangla, discussed and analysed in \citep{karim.2011}. Standard Bangla keeps coda clusters in words borrowed from English, but these are simplified in the Dhaka dialect in a variety of ways, including both epenthesis and deletion of one of the consonants in the cluster.

\ex. \a. RN clusters are broken up via copy epenthesis \label{banglaRN}
         \a. \textipa{hOrn} $\rightarrow$ \textipa{hOrOn} `horn'
         \b. \textipa{p\textsuperscript{h}Orm} $\rightarrow$ \textipa{p\textsuperscript{h}OrOm} `form'
         \c. \textipa{fIlm} $\rightarrow$ \textipa{fIlIm} `film'
         \z.
     \b. RT clusters resolved by deletion of [r] \label{banglaRT}
         \a. \textipa{park} $\rightarrow$ \textipa{pak} `park'
         \b. \textipa{\textrtailt Orc} $\rightarrow$ \textipa{\textrtailt Oc} `torch'
         \c. \textipa{narb\textsuperscript{h}} $\rightarrow$ \textipa{nab\textsuperscript{h}} `nerve'
         \z.
     \citep[(4,5)]{karim.2011}

From the examples in \ref{banglaRN}, we see that {\sc Max}(r) $\gg$ {\sc Dep} in general, making epenthesis the preferred way of dealing with the cluster.

\ex. {\sc *ComplexCoda} $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{hOrn}/} & {\sc *ComplexCoda} & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{hOrn} & $\ast$! & & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{hOrOn} & & & \cellcolor{lightgray}$\ast$ \\
\hline c. & & \textipa{hOn} & & $\ast$! & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

However, if we retain this ranking, we incorrectly expect RT clusters to be broken up by epenthesis rather than delete the [r]. This can be resolved by ranking {\sc Ident(SonAngle)}$<$2.3, say, above {\sc Max}(r). This then blocks epenthesis in RT clusters, which have a {\sc Sonority Angle} of 2.35:

\ex. {\sc *ComplexCoda}, {\sc Ident(SonAngle)}$<$2.3  $\gg$ {\sc Max}(r) $\gg$ {\sc Dep}

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c:c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{park}/} & {\sc *ComplexCoda} & {\sc Ident(SonAngle)}$<$2.3 & {\sc Max}(r) & {\sc Dep} \\[0.5ex]
\hline \hline a. & & \textipa{park} & $\ast$! & & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & & \textipa{parak} & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray}$\ast$ \\
\hline c. & \ding{43} & \textipa{pak} & & & \cellcolor{lightgray}$\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Since RN clusters have a smaller {\sc Sonority Angle} of 1.9, their behaviour does not change with this ranking.

While other analyses could possibly account for this data (see one such in \citep{karim.2011}), we see that simply ranking {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{F}$ allows us to account for it quite straightforwardly.

Thus far, we have not made any particular use of free ranking. While ranking {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{M}$ does conflate all clusters with a {\sc Sonority Angle} above $n$, as well as all clusters with a {\sc Sonority Angle} below $n$, preserving only the distinction between {\sc Sonority Angles} greater and less than $n$, this could be accomplished using a universal ranking of the {\sc Ident(SonAngle)}$<$n constraints as well, as illustrated in the following example:

\ex. The {\sc Ident(SonAngle)}$<$n with the smallest $n$ that outranks {\sc *ComplexCoda} determines the cut-off point.

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/TF/} & {\sc Id(SA)}$<$2.0 & {\sc Id(SA)}$<$1.5 & {\sc *CompCoda} & {\sc Id(SA)}$<$1.0 & {\sc Id(SA)}$<$0.5 \\[0.5ex]
\hline \hline a. & & TF & & & $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline b. & \ding{43} & T\textipa{@}F & 0.59 & 0.59 & & \cellcolor{lightgray}0.59 & \cellcolor{lightgray}0.59 $\ast$ \\
\hline \end{tabular}} %\renewcommand*\arraystretch{1} \end{center}

\bigskip
%\begin{center} \renewcommand*\arraystretch{1.2}
%\scalebox{1}[1]
{\begin{tabular}[t]{|rrl||c|c|c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/FT/} & {\sc Id(SA)}$<$2.5 & {\sc Id(SA)}$<$1.5 & {\sc *CompCoda} & {\sc Id(SA)}$<$1.0 & {\sc Id(SA)}$<$0.5 \\[0.5ex]
\hline \hline c. & \ding{43}& FT & & & \cellcolor{lightgray} $\ast$ & \cellcolor{lightgray} & \cellcolor{lightgray} \\
\hline d. &  & F\textipa{@}T & 2.11 & 2.11 $\ast$! & \cellcolor{lightgray} & \cellcolor{lightgray} 2.11 $\ast$ & \cellcolor{lightgray}2.11 $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Where free ranking comes into play is when two output candidates break up different clusters. We will see examples of this in the analysis of Chaha three- and four-consonant clusters. I will give a schematic example here to illustrate the phenomenon.

Suppose a triconsonantal cluster /CCC/ has to be broken up either as [C\textipa{@}CC] or [CC\textipa{@}C], and further that [C\textipa{@}CC] violates some additional markedness constraint $\mathcal{M}$. Which {\sc Ident(SonAngle)}$<$n constraint outranks $\mathcal{M}$ will then determine the positioning of the epenthetic vowel.

\ex. {\sc Ident(SonAngle)}$<$n $\gg$ $\mathcal{M}$
     \a. Scenario 1: {\sc SonAngle}(C1-C2), {\sc SonAngle}(C2,C3)$<$n \\
         Since neither violates {\sc Ident(SonAngle)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C1C2C3/} & {\sc Ident(SonAngle)}$<$n & $\mathcal{M}$ \\[0.5ex]
\hline \hline a. & \ding{43} & C1C2\textipa{@}C3 & $<$n & \\
\hline b. & & C1\textipa{@}C2C3 & $<$n & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
      \b. Scenario 2: {\sc SonAngle}(C1-C2)$>$n, {\sc SonAngle}(C2,C3)$<$n \\
          C1\textipa{@}C2C3 harmonically bound by C1C2\textipa{@}C3 candidate.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C1C2C3/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline c. & \ding{43} &  C1C2\textipa{@}C3  & $<$n & \cellcolor{lightgray} \\
\hline d. & & C1\textipa{@}C2C3 & $>$n $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Scenario 3: {\sc SonAngle}(C1-C2)$<$n, {\sc SonAngle}(C2,C3)$>$n \\
         {\sc Ident(SonAngle)}$<$n determines output, overruling $\mathcal{M}$.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C1C2C3/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline e. & &  C1C2\textipa{@}C3 & $>$n $\ast$!  & \cellcolor{lightgray} \\
\hline f. & \ding{43} & C1\textipa{@}C2C3& $>$n & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. Scenario 4: {\sc SonAngle}(C1-C2), {\sc SonAngle}(C2-C3) $>$ n \\
         Since both candidates violate {\sc Ident(SonAngle)}$<$n, $\mathcal{M}$ determines output.
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C1C2C3/} & {\sc Ident(SonAngle)}$<$n &$ \mathcal{M}$  \\[0.5ex]
\hline \hline g. & &  C1C2\textipa{@}C3 & $>$n $\ast$  & \\
\hline h. & \ding{43} & C1\textipa{@}C2C3 & $>$n $\ast$ & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Because we have assumed that all other {\sc Ident(SonAngle)}$<$n constraints rank low, certainly below $\mathcal{M}$, the degree to which candidates (g) and (h) violate {\sc Ident(SonAngle)}$<$n is irrelevant --- the distinction between clusters with {\sc Sonority Angles} above $n$ has been conflated. Even if {\sc Sonority Angle}(C2,C3) exceeds {\sc Sonority Angle}(C1,C2), the outcome is determined by the markedness constraint $\mathcal{M}$.

Contrast this with the universal ranking, under which this distinction would be relevant. Numbers are given for clarity.

\ex. Scenario: {\sc SonAngle}(C2-C3) $>$ {\sc SonAngle}(C1-C2) $>$ n \\

\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/C1C2C3/} & {\sc Ident(SonAngle)}$<$1.9 & {\sc Ident(SonAngle)}$<$1.5 &$ \mathcal{M}$  \\[0.5ex]
\hline \hline a. & \ding{43} &  C1C2\textipa{@}C3 & 2.0 $\ast$!  & 2.0 $\ast$ &  \\
\hline b.  & & C1\textipa{@}C2C3 & 1.7  & 1.7 $\ast$ & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

Under a universal ranking of {\sc Ident(SonAngle)}$<$n constraints, [C1C2\textipa{@}C3] would win in this scenario. In an equivalent scenario under free-ranking, where {\sc Ident(SonAngle)}$<$1.5 is the least stringent constraint to outrank $\mathcal{M}$, which in turn outranks all more stringent {\sc Ident(SonAngle)}$<$n constraints, the [C1C2\textipa{@}C3] candidate would win instead. There would be conflation of all clusters with {\sc Sonority Angles} below $n$, but distinctions between clusters with {\sc Sonority Angles} would still be relevant.

The example of Chaha will show that free-ranking is necessary, as conflation is observed on both sides of $n$, where {\sc Ident(SonAngle)}$<$n is the most stringent constraint to outrank $\mathcal{M}$.

@@

I now turn to an alternative metric of sonority contour faithfulness proposed by \citet{flemming.2008}, {\sc Sonority Rise}, which takes the ratio of the underlying sonority contour /C\textsubscript{1}C\textsubscript{2}/ and the surface contour [C\textsubscript{1}V] and subtracts that from 1.

@@

{\sc Sonority Rise} also gives rise to a hierarchy of susceptibility of clusters.

@@

Observe that of the two desiderata for such a metric, {\sc Sonority Rise} fulfills the first but not the second. Consider the falling sonority clusters. RN, NF and FT, which all have a sonority distance of -1, have a smaller {\sc Sonority Rise} than NT and RF, with a sonority distance of -2, followed by RT with a sonority distance of -3. %TODO: check if these should be + or -

However, of the three, RN has the largest {\sc Sonority Rise}, and therefore is expected to be more resistant to epenthesis than NF and FT. If RN undergoes epenthesis, therefore, we predict FT and NF to do the same.

Another prediction is that if NT and RT fail to undergo epenthesis due to the high faithfulness cost of interrupting these clusters, then RF should also fail to undergo epenthesis. Both of these predictions are contrary to the evidence of Irish and Chaha.

@@

\ex. Universal ranking cannot derive [\textipa{jams1to}]

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j-a-mst-o}/} & {\sc Ident(SonAngle)}$<$2.1 & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a.  &\frownie & \textipa{jams1to} & 2.11 $\ast$! & \cellcolor{lightgray}2.11 $\ast$ & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{jam1sto} & 2.03 & \cellcolor{lightgray}2.03 $\ast$ & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

Rose also uses the {\sc Jussive-R} constraint in her analysis of quadriconsonantal clusters.

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Align-Left}
     \a. Case 1: {\sc Sonority Angle}(C2,C3)$<$1.5
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-gdf-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{n1gd1fn@} & 0.59 & \\
\hline b. & & \textipa{n1g1df1n@} & 1.37, 0.54 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Case 2: {\sc Sonority Angle}(C2,C3)$>=$1.5 but so is one of {\sc Sonority Angle}(C1,C2) or {\sc Sonority Angle}(C3,C4)
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-srB-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{n1sr1Bn@} & 1.57 $\ast$ &  \\
\hline b. & & \textipa{n1s1rB1n@} & 0.22, 1.65$\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
      \c. Case 3: {\sc Sonority Angle}(C2,C3)$>=$1.5 and both {\sc Sonority Angle}(C1,C2) and {\sc Sonority Angle}(C3,C4)$<$1.5
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-dmd-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. &  & \textipa{n1dm1dn@} & 2.3 $\ast$!   & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{n1d1md1n@} & 0.27, 0.27 & \cellcolor{lightgray} $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Anchor-Left Rt}
     \a. Case 1: Both {\sc Sonority Angle}(C1,C2) and {\sc Sonority Angle}(C3,C4)$<$1.5
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-gdf-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. &  & \textipa{n1gd1fn@} & 0.59 & $\ast$!\\
\hline b. & \ding{43} & \textipa{n1g1df1n@} & 1.37, 0.54 &  \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. Case 2: {\sc Sonority Angle}(C1,C2) or {\sc Sonority Angle}(C3,C4) $>=$1.5, but so is {\sc Sonority Angle}(C2,C3)
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-srB-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. &  & \textipa{n1sr1Bn@} & 1.57 $\ast$ & $\ast$! \\
\hline b. & \ding{43} & \textipa{n1s1rB1n@} & 0.22, 1.65$\ast$ &  \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Case 3: {\sc Sonority Angle}(C1,C2) or {\sc Sonority Angle}(C3,C4) $>=$1.5, and {\sc Sonority Angle}(C2,C3)$<$1.5
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{n-sdB-n@}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-Left Rt} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{n1sd1Bn@} & 0.18 & \cellcolor{lightgray} $\ast$ \\
\hline b. & & \textipa{n1s1dB1n@} & 2.11 $\ast$!, 1.65 $\ast$ & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. Table of CCCC clusters, observed outcomes and predictions made by the constraint rankings proposed for the four idiolects in Section ???
